\pkg{DNest4}: Diffusive Nested Sampling in \proglang{C++} and \proglang{Python}

### 2018-08-07 paper/ms.tex
Working on replication - still needs new figures and numbers for galaxyfield and abc
Numba

The \proglang{Python}
packages \pkg{NumPy} \citep{numpy}, \pkg{matplotlib} \citep{matplotlib},
and \pkg{Cython} \citep{cython} are also needed.
To download and compile \pkg{DNest4}
the following steps are sufficient:
%
\begin{CodeChunk}
\begin{CodeInput}
$ wget https://github.com/eggplantbren/DNest4/archive/0.1.4.tar.gz
$ tar xvzf 0.1.4.tar.gz
$ mv DNest4-0.1.4 DNest4
$ cd DNest4/code
$ make
$ cd ../python
$ python setup.py install
\end{CodeInput}
\end{CodeChunk}
%
In Mac OS X, the final line (which installs the \proglang{Python} parts
of \pkg{DNest4})
needs to provide information about
your OS version. For example, if your computer runs Mac OS X 10.9,
the installation command for the \proglang{Python} package is:
%
\begin{CodeChunk}
\begin{CodeInput}
$ MACOSX_DEPLOYMENT_TARGET=10.9 python setup.py install
\end{CodeInput}
\end{CodeChunk}
%
We recommend to create an environment variable called \code{DNEST4\_PATH}
and set it to the directory {\em above} the \pkg{DNest4} directory. Then,
if the model templates from the \pkg{DNest4} repository are copied
to any other location on your system and used as the basis for
new work, their Makefiles will continue to function.

@@

\section{Postprocessing}\label{sec:postprocessing}
The output files themselves are typically not immediately useful.
The goal of running
nested sampling is usually to obtain posterior samples and the marginal likelihood $Z$,
whereas \code{sample.txt} only contains samples from the mixture of constrained
priors. Additional
post-processing is required. This can be achieved by running the following
\proglang{Python} function\footnote{Alternatively, the file
\code{showresults.py} in the example directory runs this function, and then
calls the code in \code{display.py} to create a further plot shown in
Figure~\ref{fig:fig3} on the right.}:
%
\begin{CodeChunk}
\begin{CodeInput}
>>> import dnest4
>>> dnest4.postprocess()
\end{CodeInput}
\end{CodeChunk}
%
This produces the three diagnostic plots in Figures~\ref{fig:fig1}
and~\ref{fig:fig3} on the left, along with the following output:
%
\begin{CodeChunk}
\begin{CodeOutput}
log(Z) = -175.512920079
Information = 15.1683473499 nats.
Effective sample size = 1521.58236273
\end{CodeOutput}
\end{CodeChunk}
%
These are the natural log of the marginal likelihood, the
information
\begin{align*}
\mathcal{H} &= \int p(\theta|D, M)
\ln\left[\frac{p(\theta | D, M)}{p(\theta | M)}\right] \, d\theta,
\end{align*}
which quantifies the degree to which $D$ restricted the
range of possible $\theta$ values,
and the effective sample size,
or number of saved particles with significant posterior weight.
The \code{postprocess} function also saves a file,
\code{posterior\_sample.txt}, containing posterior samples (one per row).
Unfortunately, it is harder to compute justified error bars on $\ln(Z)$
in DNS than it is in standard nested sampling.
%The \code{postprocess} function does include some options
%for calculating error bars, but they are not used by default,
%and we do not describe them here.
%Hence, the zero-sized error bars on $\ln(Z)$ and $\mathcal{H}$ should
%be ignored.

>>>>

The \proglang{Python}
packages \pkg{NumPy} \citep{numpy}, \pkg{matplotlib} \citep{matplotlib},
\pkg{numba} \citep{numba},
and \pkg{Cython} \citep{cython} are also needed.
To download and compile \pkg{DNest4}
the following steps are sufficient:
%
\begin{CodeChunk}
\begin{CodeInput}
$ wget https://github.com/eggplantbren/DNest4/archive/0.1.4.tar.gz
$ tar xvzf 0.1.4.tar.gz
$ mv DNest4-0.1.4 DNest4
$ cd DNest4/code
$ make
$ cd ../python
$ python setup.py install
\end{CodeInput}
\end{CodeChunk}
%
In Mac OS X, the final line (which installs the \proglang{Python} parts
of \pkg{DNest4})
needs to provide information about
your OS version. For example, if your computer runs Mac OS X 10.9,
the installation command for the \proglang{Python} package is:
%
\begin{CodeChunk}
\begin{CodeInput}
$ MACOSX_DEPLOYMENT_TARGET=10.9 python setup.py install
\end{CodeInput}
\end{CodeChunk}
%
We recommend to create an environment variable called \code{DNEST4\_PATH}
and set it to the directory {\em above} the \pkg{DNest4} directory. Then,
if the model templates from the \pkg{DNest4} repository are copied
to any other location on your system and used as the basis for
new work, their Makefiles will continue to function.

@@

\section{Postprocessing}\label{sec:postprocessing}
The output files themselves are typically not immediately useful.
The goal of running
nested sampling is usually to obtain posterior samples and the marginal likelihood $Z$,
whereas \code{sample.txt} only contains samples from the mixture of constrained
priors. Additional
post-processing is required. This can be achieved by running the following
\proglang{Python} function\footnote{Alternatively, the file
\code{showresults.py} in the example directory runs this function, and then
calls the code in \code{display.py} to create a further plot shown in
Figure~\ref{fig:fig3} on the right.}:
%
\begin{CodeChunk}
\begin{CodeInput}
>>> import dnest4
>>> dnest4.postprocess()
\end{CodeInput}
\end{CodeChunk}
%
This produces the three diagnostic plots in Figures~\ref{fig:fig1}
and~\ref{fig:fig3} on the left, along with the following output:
%
\begin{CodeChunk}
\begin{CodeOutput}
log(Z) = -175.49577676640516
Information = 15.17486407892028 nats.
Effective sample size = 1126.5312086470944
\end{CodeOutput}
\end{CodeChunk}
%
These are the natural log of the marginal likelihood, the
information
\begin{align*}
\mathcal{H} &= \int p(\theta|D, M)
\ln\left[\frac{p(\theta | D, M)}{p(\theta | M)}\right] \, d\theta,
\end{align*}
which quantifies the degree to which $D$ restricted the
range of possible $\theta$ values,
and the effective sample size,
or number of saved particles with significant posterior weight.
The \code{postprocess} function also saves a file,
\code{posterior\_sample.txt}, containing posterior samples (one per row).
Unfortunately, it is harder to compute justified error bars on $\ln(Z)$
in DNS than it is in standard nested sampling.
%The \code{postprocess} function does include some options
%for calculating error bars, but they are not used by default,
%and we do not describe them here.
%Hence, the zero-sized error bars on $\ln(Z)$ and $\mathcal{H}$ should
%be ignored.
